---
title: "HAR classification project"
author: "Ramón Girona"
date: "27 de diciembre de 2015"
output: html_document
---

## Overview

The objective is to perform an analysis of data from Human Activity Recognition project (HAR) corresponding to 6 different practicioners exercising Weight Lifting. From this analysis we may produce a predictor function, which can serve to classify exercises out of sample i.e. out of the training data set.  


## HAR dataset

There is a train data set and a validation data set.  highlighting basic features of the data, by using confidence intervals and/or tests.


## Methodology
The first exploratory analysis of the data show the following conclussions:


When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r}
summary(cars)
```

You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
