---
title: "HAR classification project"
author: "Ramón Girona"
date: "27 de diciembre de 2015"
output: html_document
---

# Overview

The objective is to perform an analysis of data from Human Activity Recognition project (HAR) corresponding to 6 different practicioners exercising Weight Lifting. From this analysis we may produce a predictor function, which can serve to classify exercises out of sample i.e. out of the training data set.  


# HAR dataset

There is a training data set ("pml-training.csv"), which will be used  for both training and testing, except for the k-fold cross validation, and a validation data set ("pml-testing.csv"). 

The training data set contains 2 different types of register:
- Register type 1: These are samples taken from different sensors, mainly from 4 different ones wearables: Belt, arm and forearm, and one on the dumbbell. Typeical measures taken are: gyros, accelerometer, magnetometer x,y and z for the 4 different sensors, and the corresponding translation into roll, pitch and yaw movements. There are 55 variables. 

- Register type 2: These are statistics from some of the previous measures, taken during a time window. This information comes from another field/variable num_window, and only when there is a change of window (where another field, new_window changes to "yes"), the last register contains stats like Kurtosis, Skewness, Std deviation / Variance, Average (4 moments), Min,  Max and range (amplitude) for roll, pitch, yaw and acceleration of the 4 sensors. There are 98 stat variables.   

### This doesn't apply to the testing data set where we find only registers type 1. 

# Methodology

- Define your error rate. WE will consider for this project 5%.
- Step 1: Read the data and split data into:
    Training (60% of "pml-training.csv"), Testing (the remaining 40%), Validation ("pml-testing.csv")
- Step 2: On the training set pick a prediction function (2 for the 2 dofferent register types). These will be our first tries.
- Step 3: Use cross-validation with the testing partition of the training set, using the k-fold method.
- Step 4: Apply to the validation sample ("pml-testing.csv"")

# Steps

## Step 1
The first exploratory analysis of the data show the following conclussions:
The new_window variable eq. "yes" signals where the series of measures corresponding to a window are summarized by means of average, variance, skewness, etc.

We have used the data from the registers type 2 to create our first predictor, though it only applies to registers of this type. There are only 406 registers type 2 in the training set, out of over 19000 registers. Neither of the registers of the validation set is register type 2, we may therefore produce a different kind of predictors for the latter.

We are not using predictors for one kind of register  for the other type. First we will create the 2 training and testing sets corresponding to the 2 different register types.

```{r set-up}

# Load the csv files 

HAR_fulldata <- read.csv("pml-training.csv")
HAR_testdata <- read.csv("pml-testing.csv")

# Select only the proper columns

HAR_fd_stat <- subset(HAR_fulldata[,c(12:36,50:59,69:83,87:88,90:91,93:101,103:112,125:129,131:139,141:150,160)], HAR_fulldata$new_window %in% "yes")
HAR_fd_smpls <- subset(HAR_fulldata[,c(8:11,46:49,84:86,102,122:124,140,160)], HAR_fulldata$new_window %in% "no")

# Use the caret lib

library("caret", lib.loc="~/R/x86_64-pc-linux-gnu-library/3.2")

# Partitioning the data in the training and testing sets

inTr_sm <- createDataPartition(HAR_fd_smpls$classe,p=0.6, list=FALSE)
inTr_st <- createDataPartition(HAR_fd_stat$classe,p=0.8, list=FALSE)

# _smpls suffix means register type 1
# _stat suffix means register type 2

train_smpls <- HAR_fd_smpls[inTr_sm,]
test_smpls <- HAR_fd_smpls[-inTr_sm,]
test_stat <- HAR_fd_stat[-inTr_st,]
train_stat <- HAR_fd_stat[inTr_st,]

```

## Step 2

We will make use of the 2 different kind of predictors for the 2 different type of registers.

model_fit01, used for registers type 2, while model_fit02 is intended for registers type 1. We will make the corresponding prediction and obtain the confusion matrix to check the validity of the  prediction.

For the model_fit02 we will make use of a subset of the variables, consisting in:

total_acceleration, roll, pitch and yaw applied to: Belt, arm, forearm and dumbbells. 12 variables in total. 


```{r modelling, eval=FALSE}

# model_fit01 took over an hour with so many variables. No simulation included

model_fit01 <- train(classe ~ ., data=train_stat)
model_fit02 <- train(classe ~ ., data=train_smpls, method="rf")

# prediction and test accuracy through the confussion matrix

prediction01 <- predict(model_fit01, newdata=test_stat)
confusionMatrix(prediction01,test_stat$classe)

prediction02 <- predict(model_fit02, newdata=test_smpls)
confusionMatrix(prediction02,test_smpls$classe)

```

For the first model the confusion matrix show us an accuracy of 0.83 and kappa of 0.79

Confusion Matrix and Statistics

          Reference
Prediction  A  B  C  D  E
         A 20  3  2  1  0
         B  0 11  3  0  1
         C  1  0  9  0  0
         D  0  0  0 11  0
         E  0  1  0  1 14

Overall Statistics
                                          
               Accuracy : 0.8333          
                 95% CI : (0.7319, 0.9082)
    No Information Rate : 0.2692          
    P-Value [Acc > NIR] : < 2.2e-16       
                                          
                  Kappa : 0.7883          
 Mcnemar's Test P-Value : NA              

Statistics by Class:

                     Class: A Class: B Class: C Class: D Class: E
Sensitivity            0.9524   0.7333   0.6429   0.8462   0.9333
Specificity            0.8947   0.9365   0.9844   1.0000   0.9683

For the second model, we have obtained better results, close to 100% with 0.98 of accuracy, and sensitivity and specificity over 0.97 for the 5 levels. This model is also applicable to most of the registers, included the validation ones.

          Reference
Prediction    A    B    C    D    E
         A 2178   13    0    0    0
         B    7 1446   14    0    3
         C    3   26 1307   10    0
         D    0    2   18 1246    0
         E    0    0    1    2 1408

Overall Statistics
                                          
               Accuracy : 0.9871          
                 95% CI : (0.9843, 0.9895)
    No Information Rate : 0.2847          
    P-Value [Acc > NIR] : < 2.2e-16       
                                          
                  Kappa : 0.9837          
 Mcnemar's Test P-Value : NA              

Statistics by Class:

                     Class: A Class: B Class: C Class: D Class: E
Sensitivity            0.9954   0.9724   0.9754   0.9905   0.9979
Specificity            0.9976   0.9961   0.9939   0.9969   0.9995

## Step 3

Now we will use cross-validation to prove the capability of the model oyt of the sample. For that purpose wi will make use of 3 folds method. 

The k-fold cross validation method involves splitting the dataset into k-subsets. For each subset is held out while the model is trained on all other subsets. This process is completed until accuracy is determine for each instance in the dataset, and an overall accuracy estimate is provided.

For the control of the process we will employ the traincontrol function of the caret library.

```{r cross-validation}

# Use the caret lib
library("caret", lib.loc="~/R/x86_64-pc-linux-gnu-library/3.2")

# define training control
train_control <- trainControl(method="cv", number=3)

# train the model 
model <- train(classe~., data=train_smpls, trControl=train_control, method="rf")

# make predictions
predictions <- predict(model,test_smpls)

# summarize results
confusionMatrix(predictions, test_smpls$classe)

```

# Step 4

Performed for the subsequent exercise


# Results / Conclusions

If the results are evaluated by the accueracy of the model with the validation sample, the confidence would be 100%: 20 out of 20 well predicted.

Only one caveat the validation sample is in-sample. It is well noticed that the num_window correspond always to registers existing in the training set. We checked the timestamps matched also the corresponding values.